# Inference on Predictive and Causal Effects in High-Dimensional Nonlinear Models

## Impact of 401(k) on  Financial Wealth

As a practical illustration of the methods developed in this lecture, we consider estimation of the effect of 401(k) eligibility and participation 

One can argue that eligibility for enrolling in a 401(k) plan in this data can be taken as exogenous after conditioning on a few observables of which the most important for their argument is income. The basic idea is that, at least around the time 401(k)’s initially became available, people were unlikely to be basing their employment decisions on whether an employer offered a 401(k) but would instead focus on income and other aspects of the job. 

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>

### Data

The data set can be loaded from the `hdm` package for R and hdmpy for Python by typing

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE }
library(hdm) 
library(ggplot2)
data(pension)
data <- pension 
dim(data)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
import numpy as np
import pandas as pd
from doubleml.datasets import fetch_401K
from sklearn.preprocessing import PolynomialFeatures
import matplotlib.pyplot as plt
import seaborn as sns
import hdmpy
import pyreadr
import urllib.request
import os
import warnings
warnings.filterwarnings("ignore")

rdata_read = pyreadr.read_r("./data/pension.Rdata")
# Extracting the data frame from rdata_read
data = rdata_read[ 'data' ]
pension = data.copy()
data.shape

```
:::
::::::
\newline

See the "Details" section on the description of the data set, which can be accessed by 

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
help(pension)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
os.system("start \"\" https://search.r-project.org/CRAN/refmans/hdm/html/pension.html")
```
:::
::::::
\newline


The data consist of 9,915 observations at the household level drawn from the 1991 Survey of Income and Program Participation (SIPP).  All the variables are referred to 1990. We use net financial assets (*net\_tfa*) as the outcome variable, $Y$,  in our analysis. The net financial assets are computed as the sum of IRA balances, 401(k) balances, checking accounts, saving bonds, other interest-earning accounts, other interest-earning assets, stocks, and mutual funds less non mortgage debts. 

Among the $9915$ individuals, $3682$ are eligible to participate in the program. The variable *e401* indicates eligibility and *p401* indicates participation, respectively.


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
hist_e401 = ggplot(data, aes(x = e401, fill = factor(e401))) + 
            geom_bar()
hist_e401
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
sns.set()
colors = sns.color_palette()
sns.countplot(x="e401", hue = "e401" , data=data)
```
:::
::::::
\newline





Eligibility is highly associated with financial wealth:


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE, fig.height=5, fig.width=15}
dens_net_tfa = ggplot(data, aes(x = net_tfa, color = factor(e401), fill = factor(e401)) ) + 
                    geom_density() + xlim(c(-20000, 150000)) + 
                    facet_wrap(.~e401) 
dens_net_tfa
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}
from matplotlib.pyplot import figure
figure(figsize=(8, 10), dpi=100)
sns.displot(data, x="net_tfa", hue="e401", col="e401",
                kind="kde", fill=True)
```
:::
::::::
\newline











The unconditional APE of e401 is about $19559$:

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
e1 <- data[data$e401==1,]
e0 <- data[data$e401==0,]
round(mean(e1$net_tfa)-mean(e0$net_tfa),0)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
int(np.round( data[['e401', 'net_tfa']].groupby('e401').mean().diff().iloc[1, 0] ) )
```
:::
::::::
\newline














Among the $3682$ individuals that  are eligible, $2594$ decided to participate in the program. The unconditional APE of p401 is about $27372$:


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
p1 <- data[data$p401==1,]
p0 <- data[data$p401==0,]
round(mean(p1$net_tfa)-mean(p0$net_tfa),0)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
int(np.round( data[['p401', 'net_tfa']].groupby('p401').mean().diff().iloc[1, 0]))
```
:::
::::::
\newline













As discussed, these estimates are biased since they do not account for saver heterogeneity and endogeneity of participation.

## Double ML package

We are interested in valid estimators of the average treatment effect of `e401` and `p401` on `net_tfa`. To get those estimators, we use the `DoubleML` package that internally builds on mlr3. You find additional information on the package on the package website https://docs.doubleml.org/ and the R documentation page https://docs.doubleml.org/r/stable/. 


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# installing Double ML
#remotes::install_github("DoubleML/doubleml-for-r",quiet=TRUE)


# loading the packages
library(DoubleML)
library(mlr3learners)
library(mlr3)
library(data.table)
library(randomForest)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
import doubleml as dml
from sklearn.linear_model import LassoCV, LogisticRegressionCV
from sklearn.ensemble import RandomForestClassifier, RandomForestRegressor
from sklearn.tree import DecisionTreeClassifier, DecisionTreeRegressor
from sklearn.preprocessing import StandardScaler
from sklearn.pipeline import make_pipeline
from xgboost import XGBClassifier, XGBRegressor
```
:::
::::::
\newline













As mentioned, in the tutorial we use the meta package `mlr3` to generate predictions with machine learning methods. A comprehensive introduction and description of the `mlr3` package is provided in the [mlr3book](https://mlr3book.mlr-org.com/). A list of all learners that you can use in `mlr3` can be found [here](https://mlr3extralearners.mlr-org.com/articles/learners/list_learners.html). The entry in the columns *mlr3 Package* and *Packages* indicate which packages must be installed/loaded in your R session. 

## Estimating the ATE of 401(k) Eligibility on Net Financial Assets

We first look at the treatment effect of e401 on net total financial assets. We give estimates of the ATE and ATT that corresponds to the linear model

\begin{equation*}
Y = D \alpha + f(X)'\beta+ \epsilon,
\end{equation*}

where $f(X)$ includes indicators of marital status, two-earner status, defined benefit pension status, IRA participation status, and home ownership status, and  orthogonal polynomials of degrees 2, 4, 6 and 8 in family size, education, age and  income, respectively. The dimensions of $f(X)$ is 25. 

In the first step, we report estimates of the average treatment effect (ATE) of 401(k) eligibility on net financial assets both in the partially linear regression (PLR) model and in the interactive regression model (IRM) allowing for heterogeneous treatment effects. 

I decided to low down the degree of the income variable from **8**  to **6** since sklearn has some problems with values greater than `float32`. I can not use `np.log` in `inc` variable since it has negative values. As a result, we got a column with some `None` values. For more information about this problem, check this [link](https://github.com/scikit-learn/scikit-learn/issues/15628).

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# Constructing the data (as DoubleMLData)

formula_flex = "net_tfa ~ e401 + poly(age, 6, raw=TRUE) + poly(inc, 6, raw=TRUE) + poly(educ, 4, raw=TRUE) + poly(fsize, 2, raw=TRUE) + marr + twoearn + db + pira + hown"

model_flex = as.data.table(model.frame(formula_flex, pension))


x_cols = colnames(model_flex)[-c(1,2)]


data_ml_aux = DoubleMLData$new(model_flex, 
                               y_col = "net_tfa", 
                               d_cols = "e401", 
                               x_cols=x_cols)


p <- dim(model_flex)[2]-2

p


# complex model with two-way interactions
#data_interactions = fetch_401k(polynomial_features = TRUE, instrument = FALSE)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Constructing the data (as DoubleMLData)
features = data.copy()[['marr', 'twoearn', 'db', 'pira', 'hown']]

poly_dict = {'age': 6,
             'inc': 6,
             'educ': 4,
             'fsize': 2}
for key, degree in poly_dict.items():
    poly = PolynomialFeatures(degree, include_bias=False)
    data_transf = poly.fit_transform(data[[key]])
    x_cols = poly.get_feature_names([key])
    data_transf = pd.DataFrame(data_transf, columns=x_cols)
    
    features = pd.concat((features, data_transf),
                          axis=1, sort=False)

model_flex = pd.concat((data.copy()[['net_tfa', 'e401']], features.copy()),
                        axis=1, sort=False)

x_cols = model_flex.columns.to_list()[2:]

# Initialize DoubleMLData (data-backend of DoubleML)
data_ml_aux = dml.DoubleMLData(model_flex, y_col='net_tfa', \
                           d_cols ='e401' , x_cols = x_cols)

# complex model with two-way interactions
# data_interactions
# fetch_401K( return_type  = 'DataFrame' , polynomial_features = True )

p = model_flex.shape[1] - 2
print(p)

p = model_flex.shape[1] - 2
print(p)
```
:::
::::::
\newline












## Partially Linear Regression Models (PLR)

We start using lasso to estimate the function $g_0$ and $m_0$ in the following PLR model:

\begin{eqnarray}
 &  Y = D\theta_0 + g_0(X) + \zeta,  &  E[\zeta \mid D,X]= 0,\\
 & D = m_0(X) +  V,   &  E[V \mid X] = 0.
\end{eqnarray}

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# Estimating the PLR
lgr::get_logger("mlr3")$set_threshold("warn") 
set.seed(123)
lasso <- lrn("regr.cv_glmnet",nfolds = 5, s = "lambda.min")
lasso_class <- lrn("classif.cv_glmnet", nfolds = 5, s = "lambda.min")

dml_plr <- DoubleMLPLR$new(data_ml_aux, ml_g = lasso, ml_m = lasso_class, n_folds=3)
dml_plr$fit(store_predictions=TRUE)
dml_plr$summary()
lasso_plr <- dml_plr$coef
lasso_std_plr <- dml_plr$se
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Estimating the PLR
# Initialize learners
Cs = 0.0001*np.logspace(0, 4, 10)
lasso = make_pipeline(StandardScaler(), LassoCV(cv=5, max_iter=10000))
lasso_class = make_pipeline(StandardScaler(),
                            LogisticRegressionCV(cv=5, penalty='l1', solver='liblinear',
                                                 Cs = Cs, max_iter=1000))

np.random.seed(123)
# Initialize DoubleMLPLR model
dml_plr = dml.DoubleMLPLR(data_ml_aux,
                                ml_g = lasso,
                                ml_m = lasso_class,
                                n_folds = 3)

dml_plr.fit(store_predictions=True)

lasso_plr = dml_plr.summary.coef[0]
lasso_std_plr = dml_plr.summary['std err'][0]
dml_plr.summary
```
:::
::::::
\newline





Let us check the predictive performance of this model.

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
dml_plr$params_names()
g_hat <- as.matrix(dml_plr$predictions$ml_g) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
g_hat = dml_plr.predictions['ml_g'].flatten() # predictions of g_o
m_hat = dml_plr.predictions['ml_m'].flatten() # predictions of m_o
print(dml_plr.params_names)
```
:::
::::::
\newline






:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
d <- as.matrix(pension$e401) 
predictions_y <- as.matrix(d*theta)+g_hat # predictions for y
lasso_y_rmse <- sqrt(mean((y-predictions_y)^2)) 
lasso_y_rmse
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# cross-fitted RMSE: outcome
y = pension.net_tfa.to_numpy() # true observations
theta = dml_plr.coef[ 0 ] # estimated regression coefficient
d = pension.e401.to_numpy()
predictions_y = d*theta + g_hat  # predictions for y
lasso_y_rmse = np.sqrt( np.mean( ( y - predictions_y ) ** 2 ) ) 
lasso_y_rmse
```
:::
::::::
\newline



:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# cross-fitted RMSE: treatment
d <- as.matrix(pension$e401) 
lasso_d_rmse <- sqrt(mean((d-m_hat)^2)) 
lasso_d_rmse

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# cross-fitted RMSE: treatment
d = pension.e401.to_numpy()
lasso_d_rmse = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )
print( lasso_d_rmse )
# cross-fitted ce: treatment
print(np.mean( ( m_hat > 0.5 ) * 1 != d ))
```
:::
::::::
\newline







Then, we repeat this procedure for various machine learning methods.

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# Random Forest

lgr::get_logger("mlr3")$set_threshold("warn") 


randomForest <- lrn("regr.ranger")


randomForest_class <- lrn("classif.ranger")


dml_plr <- DoubleMLPLR$new(data_ml_aux, 
                           ml_g = randomForest, 
                           ml_m = randomForest_class, 
                           n_folds=3)


dml_plr$fit(store_predictions=TRUE) # set store_predictions=TRUE to evaluate the model
forest_plr <- dml_plr$coef
forest_std_plr <- dml_plr$se
dml_plr$summary()

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python message=FALSE, warning=FALSE}
# Random Forest
randomForest = RandomForestRegressor(
    n_estimators = 500  )

randomForest_class = RandomForestClassifier(
    n_estimators = 500 )


dml_plr = dml.DoubleMLPLR(data_ml_aux, ml_g = randomForest, \
                         ml_m = randomForest_class, n_folds = 3)
dml_plr.fit(store_predictions=True) # set store_predictions=TRUE to evaluate the model

forest_plr = dml_plr.coef
forest_std_plr = dml_plr.summary[ 'std err' ]

dml_plr.summary
```
:::
::::::
\newline



We can compare the accuracy of this model to the model that has been estimated with lasso.


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# Evaluation predictions
g_hat <- as.matrix(dml_plr$predictions$ml_g) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
predictions_y <- as.matrix(d*theta)+g_hat # predictions for y
forest_y_rmse <- sqrt(mean((y-predictions_y)^2)) 


# cross-fitted RMSE: treatment
forest_d_rmse <- sqrt(mean((d-m_hat)^2)) 


#### Results
forest_y_rmse
forest_d_rmse
# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python message=FALSE, warning=FALSE}
# Evaluation predictions
g_hat = dml_plr.predictions['ml_g'].flatten() # predictions of g_o
m_hat = dml_plr.predictions['ml_m'].flatten() # predictions of m_o
y = pension.net_tfa.to_numpy()
theta = dml_plr.coef[ 0 ] # estimated regression coefficient
predictions_y = d*theta + g_hat # predictions for y
forest_y_rmse = np.sqrt( np.mean( ( y - predictions_y ) ** 2 ) ) 

# cross-fitted RMSE: treatment
forest_d_rmse = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )

#### Results
print( forest_y_rmse )
print( forest_d_rmse )
# cross-fitted ce: treatment
np.mean( ( m_hat > 0.5 ) * 1 != d )
```
:::
::::::
\newline






:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# Trees
lgr::get_logger("mlr3")$set_threshold("warn") 

trees <- lrn("regr.rpart")
trees_class <- lrn("classif.rpart")

dml_plr <- DoubleMLPLR$new(data_ml_aux, ml_g = trees, ml_m = trees_class, n_folds=3)
dml_plr$fit(store_predictions=TRUE)
dml_plr$summary()
tree_plr <- dml_plr$coef
tree_std_plr <- dml_plr$se

# Evaluation predictions
g_hat <- as.matrix(dml_plr$predictions$ml_g) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
predictions_y <- as.matrix(d*theta)+g_hat # predictions for y
tree_y_rmse <- sqrt(mean((y-predictions_y)^2)) 
tree_y_rmse

# cross-fitted RMSE: treatment
tree_d_rmse <- sqrt(mean((d-m_hat)^2)) 
tree_d_rmse

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Trees
trees = DecisionTreeRegressor(
        max_depth=30, ccp_alpha=0.01, min_samples_split=20, \
        min_samples_leaf= np.round(20/3).astype(int))

trees_class = DecisionTreeClassifier( max_depth=30, ccp_alpha=0.01, \
                min_samples_split=20, \
                min_samples_leaf= np.round(20/3).astype(int) )

np.random.seed(123)
dml_plr = dml.DoubleMLPLR(data_ml_aux,
                               ml_g = trees,
                               ml_m = trees_class,
                               n_folds = 3)
dml_plr.fit(store_predictions=True)
tree_summary = dml_plr.summary
print(tree_summary)

dml_plr.fit(store_predictions=True)
dml_plr.summary

tree_plr = dml_plr.coef
tree_std_plr = dml_plr.summary[ 'std err' ]

# Evaluation predictions
g_hat = dml_plr.predictions['ml_g'].flatten() # predictions of g_o
m_hat = dml_plr.predictions['ml_m'].flatten() # predictions of m_o

y = pension.net_tfa.to_numpy()
theta = dml_plr.coef[ 0 ] # estimated regression coefficient
predictions_y = d*theta + g_hat # predictions for y
tree_y_rmse = np.sqrt( np.mean( ( y - predictions_y ) ** 2 ) ) 
print( tree_y_rmse )

# cross-fitted RMSE: treatment
tree_d_rmse = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )
print( tree_d_rmse )
# cross-fitted ce: treatment
np.mean( ( m_hat > 0.5 ) * 1 != d )
```
:::
::::::
\newline





:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# Boosting
lgr::get_logger("mlr3")$set_threshold("warn") 
boost<- lrn("regr.xgboost",objective="reg:squarederror")
boost_class <- lrn("classif.xgboost",objective = "binary:logistic",eval_metric ="logloss")

dml_plr <- DoubleMLPLR$new(data_ml_aux, ml_g = boost, ml_m = boost_class, n_folds=3)
dml_plr$fit(store_predictions=TRUE)
dml_plr$summary()
boost_plr <- dml_plr$coef
boost_std_plr <- dml_plr$se

# Evaluation predictions
g_hat <- as.matrix(dml_plr$predictions$ml_g) # predictions of g_o
m_hat <- as.matrix(dml_plr$predictions$ml_m) # predictions of m_o
theta <- as.numeric(dml_plr$coef) # estimated regression coefficient
predictions_y <- as.matrix(d*theta)+g_hat # predictions for y
boost_y_rmse <- sqrt(mean((y-predictions_y)^2)) 
boost_y_rmse

# cross-fitted RMSE: treatment
boost_d_rmse <- sqrt(mean((d-m_hat)^2)) 
boost_d_rmse

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Boosted Trees
boost = XGBRegressor(n_jobs=1, objective = "reg:squarederror" )
boost_class = XGBClassifier(use_label_encoder=False, n_jobs=1,
                            objective = "binary:logistic", \
                            eval_metric = "logloss" )

np.random.seed(123)
dml_plr = dml.DoubleMLPLR( data_ml_aux ,
                                ml_g = boost,
                                ml_m = boost_class,
                                n_folds = 3)
dml_plr.fit(store_predictions=True)
boost_summary = dml_plr.summary
print( boost_summary )

boost_plr = dml_plr.coef
boost_std_plr = dml_plr.summary[ 'std err' ]

# Evaluation predictions
g_hat = dml_plr.predictions['ml_g'].flatten() # predictions of g_o
m_hat = dml_plr.predictions['ml_m'].flatten() # predictions of m_o
y = pension.net_tfa.to_numpy()
theta = dml_plr.coef[ 0 ]  # estimated regression coefficient
predictions_y = d*theta + g_hat # predictions for y

boost_y_rmse = np.sqrt( np.mean( ( y - predictions_y ) ** 2 ) ) 
print( boost_y_rmse )

# cross-fitted RMSE: treatment
boost_d_rmse = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )
print( boost_d_rmse )

# cross-fitted ce: treatment
np.mean( ( m_hat > 0.5 ) * 1 != d )
```
:::
::::::
\newline



Let's sum up the results:

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
library(xtable)
table <- matrix(0, 4, 4)
table[1,1:4]   <- c(lasso_plr,forest_plr,tree_plr,boost_plr)
table[2,1:4]   <- c(lasso_std_plr,forest_std_plr,tree_std_plr,boost_std_plr)
table[3,1:4]   <- c(lasso_y_rmse,forest_y_rmse,tree_y_rmse,boost_y_rmse)
table[4,1:4]   <- c(lasso_d_rmse,forest_d_rmse,tree_d_rmse,boost_d_rmse)
rownames(table) <- c("Estimate","Std.Error","RMSE Y","RMSE D")
colnames(table) <- c("Lasso","Random Forest","Trees","Boosting")
tab<- xtable(table, digits = 2)
tab

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
table = np.zeros( (4, 4) )

table[0,0:4] = lasso_plr,forest_plr[0],tree_plr[0],boost_plr[0]
table[1,0:4] = lasso_std_plr,forest_std_plr,tree_std_plr,boost_std_plr
table[2,0:4] = lasso_y_rmse,forest_y_rmse,tree_y_rmse,boost_y_rmse
table[3,0:4] = lasso_d_rmse,forest_d_rmse,tree_d_rmse,boost_d_rmse
table_pd = pd.DataFrame( table , index = ["Estimate","Std.Error","RMSE Y","RMSE D" ], \
            columns = [ "Lasso","Random Forest","Trees","Boosting"])

table_pd.to_latex()
```
:::
::::::
\newline


The best model with lowest RMSE in both equation is the PLR model estimated via lasso. It gives the following estimate:

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
lasso_plr
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
lasso_plr
```
:::
::::::
\newline


## Interactive Regression Model (IRM)

Next, we consider estimation of average treatment effects when treatment effects are fully heterogeneous:

 \begin{eqnarray}\label{eq: HetPL1}
 & Y  = g_0(D, X) + U,  &  \quad E[U \mid X, D]= 0,\\
  & D  = m_0(X) + V,  & \quad  E[V\mid X] = 0.
\end{eqnarray}

To reduce the disproportionate impact of extreme propensity score weights in the interactive model
we trim the propensity scores which are close to the bounds.



:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
set.seed(123)


lgr::get_logger("mlr3")$set_threshold("warn") 


dml_irm = DoubleMLIRM$new(data_ml_aux, 
                          ml_g = lasso, 
                          ml_m = lasso_class, 
                          trimming_threshold = 0.01, 
                          n_folds=3)


dml_irm$fit(store_predictions=TRUE)


dml_irm$summary()
lasso_irm <- dml_irm$coef
lasso_std_irm <- dml_irm$se


# predictions
dml_irm$params_names()

g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d*g1_hat+(1-d)*g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_o
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Lasso
lasso = make_pipeline(StandardScaler(), LassoCV(cv=5, max_iter=20000))

# Initialize DoubleMLIRM model
np.random.seed(123)
dml_irm = dml.DoubleMLIRM(data_ml_aux, ml_g = lasso, ml_m = lasso_class, trimming_threshold = 0.01, n_folds = 3)

dml_irm.fit(store_predictions=True)

lasso_summary = dml_irm.summary
print(dml_irm.summary)
lasso_irm = dml_irm.coef[0]
lasso_std_irm = dml_irm.se[0]

# predictions
print(dml_irm.params_names)
g0_hat = dml_irm.predictions['ml_g0'].flatten() # predictions of g_0(D=0, X)
g1_hat = dml_irm.predictions['ml_g1'].flatten() # predictions of g_0(D=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
m_hat = dml_irm.predictions['ml_m'].flatten() # predictions of m_o
```
:::
::::::
\newline






:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401) 
lasso_y_irm <- sqrt(mean((y-g_hat)^2)) 
lasso_y_irm

# cross-fitted RMSE: treatment
lasso_d_irm <- sqrt(mean((d-m_hat)^2)) 
lasso_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# cross-fitted RMSE: outcome
y = pension.net_tfa.to_numpy() # true observations
d = pension.e401.to_numpy()
lasso_y_irm = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print(lasso_y_irm)

# cross-fitted RMSE: treatment
lasso_d_irm = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )
print( lasso_d_irm )

# cross-fitted ce: treatment
np.mean( ( m_hat > 0.5 ) * 1 != d ) 
```
:::
::::::
\newline



:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
##### forest #####


dml_irm = DoubleMLIRM$new(data_ml_aux, 
                          ml_g = randomForest, 
                          ml_m = randomForest_class, 
                          trimming_threshold = 0.01, 
                          n_folds=3)


dml_irm$fit(store_predictions=TRUE)

#Summary
dml_irm$summary()

forest_irm <- dml_irm$coef
forest_std_irm <- dml_plr$se

# predictions
g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d*g1_hat+(1-d)*g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401) 
forest_y_irm <- sqrt(mean((y-g_hat)^2)) 
forest_y_irm

# cross-fitted RMSE: treatment
forest_d_irm <- sqrt(mean((d-m_hat)^2)) 
forest_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)



##### trees #####


dml_irm <- DoubleMLIRM$new(data_ml_aux, 
                           ml_g = trees, 
                           ml_m = trees_class, 
                           trimming_threshold = 0.01, 
                           n_folds=3)


dml_irm$fit(store_predictions=TRUE)

dml_irm$summary()

tree_irm <- dml_irm$coef
tree_std_irm <- dml_irm$se

# predictions
g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d*g1_hat+(1-d)*g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401) 
tree_y_irm <- sqrt(mean((y-g_hat)^2)) 
tree_y_irm

# cross-fitted RMSE: treatment
tree_d_irm <- sqrt(mean((d-m_hat)^2)) 
tree_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)

##### boosting #####

dml_irm <- DoubleMLIRM$new(data_ml_aux, 
                           ml_g = boost, 
                           ml_m = boost_class,
                           trimming_threshold = 0.01, 
                           n_folds=3)

dml_irm$fit(store_predictions=TRUE)

dml_irm$summary()

boost_irm <- dml_irm$coef
boost_std_irm <- dml_irm$se

# predictions
g0_hat <- as.matrix(dml_irm$predictions$ml_g0) # predictions of g_0(D=0, X)
g1_hat <- as.matrix(dml_irm$predictions$ml_g1) # predictions of g_0(D=1, X)
g_hat <- d*g1_hat+(1-d)*g0_hat # predictions of g_0
m_hat <- as.matrix(dml_irm$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$e401) 
boost_y_irm <- sqrt(mean((y-g_hat)^2)) 
boost_y_irm

# cross-fitted RMSE: treatment
boost_d_irm <- sqrt(mean((d-m_hat)^2)) 
boost_d_irm

# cross-fitted ce: treatment
mean(ifelse(m_hat > 0.5, 1, 0) != d)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
##### forest #####
randomForest = RandomForestRegressor(n_estimators=500)
randomForest_class = RandomForestClassifier(n_estimators=500)

np.random.seed(123)
dml_irm = dml.DoubleMLIRM(data_ml_aux,
                                 ml_g = randomForest,
                                 ml_m = randomForest_class,
                                 trimming_threshold = 0.01,
                                 n_folds = 3)
dml_irm.fit( store_predictions = True ) 

print( dml_irm.summary )

forest_irm = dml_irm.coef[ 0 ]
forest_std_irm = dml_irm.se[ 0 ]

# predictions
g0_hat = dml_irm.predictions['ml_g0'].flatten() # predictions of g_0(D=0, X)
g1_hat = dml_irm.predictions['ml_g1'].flatten() # predictions of g_0(D=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
m_hat = dml_irm.predictions['ml_m'].flatten() # predictions of m_o

# cross-fitted RMSE: outcome
y = pension.net_tfa.to_numpy()
d = pension.e401.to_numpy()
forest_y_irm = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( forest_y_irm )

# cross-fitted RMSE: treatment
forest_d_irm = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )
print( forest_d_irm )

# cross-fitted ce: treatment
np.mean( ( m_hat > 0.5 ) * 1 != d )

##### trees #####
trees = DecisionTreeRegressor( max_depth=30, ccp_alpha=0.01, min_samples_split=20, min_samples_leaf=67)
trees_class = DecisionTreeClassifier( max_depth=30, ccp_alpha=0.01, min_samples_split=20, min_samples_leaf=34)
np.random.seed(123)
dml_irm = dml.DoubleMLIRM(data_ml_aux,
                                 ml_g = trees,
                                 ml_m = trees_class,
                                 trimming_threshold = 0.01,
                                 n_folds = 3)

dml_irm.fit( store_predictions = True ) 

print(dml_irm.summary)


tree_irm = dml_irm.coef[ 0 ]
tree_std_irm = dml_irm.se[ 0 ]



# predictions
g0_hat = dml_irm.predictions['ml_g0'].flatten() # predictions of g_0(D=0, X)
g1_hat = dml_irm.predictions['ml_g1'].flatten() # predictions of g_0(D=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
m_hat = dml_irm.predictions['ml_m'].flatten() # predictions of m_o

# cross-fitted RMSE: outcome
y = pension.net_tfa.to_numpy()
d = pension.e401.to_numpy()
tree_y_irm = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( tree_y_irm )

# cross-fitted RMSE: treatment
tree_d_irm = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )
print( tree_d_irm )

# cross-fitted ce: treatment
np.mean( ( m_hat > 0.5 ) * 1 != d )

##### boosting #####
boost = XGBRegressor(n_jobs=1, objective = "reg:squarederror")
boost_class = XGBClassifier(use_label_encoder=False, n_jobs=1, 
				objective = "binary:logistic", 
				eval_metric = "logloss")
np.random.seed(123)
dml_irm = dml.DoubleMLIRM(data_ml_aux, ml_g = boost, ml_m = boost_class, 
			  trimming_threshold = 0.01, n_folds = 3)
dml_irm.fit( store_predictions = True ) 

print( dml_irm.summary )


boost_irm = dml_irm.coef[ 0 ]
boost_std_irm = dml_irm.se[ 0 ]

# predictions
g0_hat = dml_irm.predictions['ml_g0'].flatten() # predictions of g_0(D=0, X)
g1_hat = dml_irm.predictions['ml_g1'].flatten() # predictions of g_0(D=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
m_hat = dml_irm.predictions['ml_m'].flatten() # predictions of m_o

# cross-fitted RMSE: outcome
y = pension.net_tfa.to_numpy()
d = pension.e401.to_numpy()
boost_y_irm = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( boost_y_irm )

# cross-fitted RMSE: treatment
boost_d_irm = np.sqrt( np.mean( ( d - m_hat ) ** 2 ) )
print( boost_d_irm )

# cross-fitted ce: treatment
np.mean( ( m_hat > 0.5 ) * 1 != d ) 
```
:::
::::::
\newline





:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
library(xtable)
table <- matrix(0, 4, 4)
table[1,1:4]   <- c(lasso_irm,forest_irm,tree_irm,boost_irm)
table[2,1:4]   <- c(lasso_std_irm,forest_std_irm,tree_std_irm,boost_std_irm)
table[3,1:4]   <- c(lasso_y_irm,forest_y_irm,tree_y_irm,boost_y_irm)
table[4,1:4]   <- c(lasso_d_irm,forest_d_irm,tree_d_irm,boost_d_irm)
rownames(table) <- c("Estimate","Std.Error","RMSE Y","RMSE D")
colnames(table) <- c("Lasso","Random Forest","Trees","Boosting")
tab<- xtable(table, digits = 2)
tab
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
table2 = np.zeros( (4, 4) )
table2[0,0:4] = lasso_irm,forest_irm,tree_irm,boost_irm
table2[1,0:4] = lasso_std_irm,forest_std_irm,tree_std_irm,boost_std_irm
table2[2,0:4] = lasso_y_irm,forest_y_irm,tree_y_irm,boost_y_irm
table2[3,0:4] = lasso_d_irm,forest_d_irm,tree_d_irm,boost_d_irm
table2_pd = pd.DataFrame( 
		table2 , 
		index = ["Estimate","Std.Error","RMSE Y","RMSE D" ], \
	        columns = [ "Lasso","Random Forest","Trees","Boosting"])
table2_pd.to_latex()
```
:::
::::::
\newline




Here, Random Forest gives the best prediction rule for $g_0$ and Lasso the best prediction rule for $m_0$, respectively. Let us fit the IRM model using the best ML method for each equation to get a final estimate for the treatment effect of eligibility.

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
set.seed(123)

lgr::get_logger("mlr3")$set_threshold("warn") 

dml_irm = DoubleMLIRM$new(data_ml_aux, 
                          ml_g = randomForest, 
                          ml_m = lasso_class, 
                          trimming_threshold = 0.01, 
                          n_folds=3)

dml_irm$fit(store_predictions=TRUE)

dml_irm$summary()
best_irm <- dml_irm$coef
best_std_irm <- dml_irm$se
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
np.random.seed(123)

dml_irm = dml.DoubleMLIRM(data_ml_aux,
                          ml_g = randomForest,
                          ml_m = lasso_class,
                          trimming_threshold = 0.01,
                          n_folds = 3)

dml_irm.fit(store_predictions=True) 


print( dml_irm.summary )
best_irm = dml_irm.coef[0]
best_std_irm = dml_irm.se[0]
```
:::
::::::
\newline





These estimates that flexibly account for confounding are
substantially attenuated relative to the baseline estimate (*19559*) that does not account for confounding. They suggest much smaller causal effects of 401(k) eligiblity on financial asset holdings. 

## Local Average Treatment Effects of 401(k) Participation on Net Financial Assets

## Interactive IV Model (IIVM)

Now, we consider estimation of local average treatment effects (LATE) of participation with the binary instrument `e401`. As before, $Y$ denotes the outcome `net_tfa`, and $X$ is the vector of covariates.  Here the structural equation model is:

\begin{eqnarray}
& Y = g_0(Z,X) + U, &\quad E[U\mid Z,X] = 0,\\
& D = r_0(Z,X) + V, &\quad E[V\mid Z, X] = 0,\\
& Z = m_0(X) + \zeta, &\quad E[\zeta \mid X] = 0.
\end{eqnarray}


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# Constructing the data (as DoubleMLData)

formula_flex2 = "net_tfa ~ p401+ e401 + 
                poly(age, 6, raw=TRUE) + 
                poly(inc, 6, raw=TRUE) + 
                poly(educ, 4, raw=TRUE) + 
                poly(fsize, 2, raw=TRUE) + 
                marr + twoearn + db + 
                pira + hown"


model_flex2 = as.data.table(model.frame(formula_flex2, data))


x_cols = colnames(model_flex2)[-c(1,2,3)]


data_IV_aux = DoubleMLData$new(model_flex2, y_col = "net_tfa", d_cols = "p401", z_cols ="e401",x_cols=x_cols)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Constructing the data (as DoubleMLData)
features = data.copy()[['marr', 'twoearn', 'db', 'pira', 'hown']]
poly_dict = {'age': 6, 'inc': 6, 'educ': 4, 'fsize': 2}
for key, degree in poly_dict.items():
    poly = PolynomialFeatures(degree, include_bias=False)
    data_transf = poly.fit_transform(data[[key]])
    x_cols = poly.get_feature_names([key])
    data_transf = pd.DataFrame(data_transf, columns=x_cols)    
    features = pd.concat((features, data_transf),
                          axis=1, sort=False)
model_flex2 = pd.concat((data.copy()[['net_tfa', 'p401' , 'e401']], features.copy()),
                        axis=1, sort=False)
x_cols = model_flex2.columns.to_list()[3:]
# Initialize DoubleMLData (data-backend of DoubleML)
data_IV_aux = dml.DoubleMLData(model_flex2, y_col='net_tfa', \
                               d_cols ='p401' , z_cols = 'e401' , \
                               x_cols = x_cols)
```
:::
::::::
\newline





:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
set.seed(123)

lgr::get_logger("mlr3")$set_threshold("warn") 

dml_MLIIVM = DoubleMLIIVM$new(data_IV_aux, 
                              ml_g = lasso, 
                              ml_m = lasso_class, 
                              ml_r = lasso_class, 
                              n_folds=3, 
                              subgroups = list(always_takers = FALSE, 
                                          never_takers = TRUE))

dml_MLIIVM$fit(store_predictions=TRUE)

dml_MLIIVM$summary()

lasso_MLIIVM <- dml_MLIIVM$coef
lasso_std_MLIIVM <- dml_MLIIVM$se
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Lasso
lasso = make_pipeline(StandardScaler(), LassoCV(cv=5, max_iter=20000))

# Initialize DoubleMLIRM model
np.random.seed(123)
dml_MLIIVM = dml.DoubleMLIIVM(data_IV_aux, 
				ml_g = lasso, 
				ml_m = lasso_class,
        ml_r = lasso_class,
        subgroups = {'always_takers': False,
                    'never_takers': True},
        trimming_threshold = 0.01,
        n_folds = 3)

dml_MLIIVM.fit(store_predictions=True)
print( dml_MLIIVM.summary )
lasso_MLIIVM = dml_MLIIVM.coef[ 0 ]
lasso_std_MLIIVM = dml_MLIIVM.se[ 0 ]

```
:::
::::::
\newline






The confidence interval for the local average treatment effect of participation is given by

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
dml_MLIIVM$confint(level = 0.95)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
dml_MLIIVM.confint()
```
:::
::::::
\newline




Here we can also check the accuracy of the model:

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# variables
y <- as.matrix(pension$net_tfa) # true observations
d <- as.matrix(pension$p401) 
z <- as.matrix(pension$e401) 

# predictions
dml_MLIIVM$params_names()
g0_hat <- as.matrix(dml_MLIIVM$predictions$ml_g0) # predictions of g_0(z=0, X)
g1_hat <- as.matrix(dml_MLIIVM$predictions$ml_g1) # predictions of g_0(z=1, X)
g_hat <- z*g1_hat+(1-z)*g0_hat # predictions of g_0
r0_hat <- as.matrix(dml_MLIIVM$predictions$ml_r0) # predictions of r_0(z=0, X)
r1_hat <- as.matrix(dml_MLIIVM$predictions$ml_r1) # predictions of r_0(z=1, X)
r_hat <- z*r1_hat+(1-z)*r0_hat # predictions of r_0
m_hat <- as.matrix(dml_MLIIVM$predictions$ml_m) # predictions of m_o
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# variables
y = pension.net_tfa.to_numpy()
d = pension.p401.to_numpy()
z = pension.e401.to_numpy()

# predictions
print( dml_MLIIVM.params_names )
g0_hat = dml_MLIIVM.predictions['ml_g0'].flatten() # predictions of g_0(z=0, X)
g1_hat = dml_MLIIVM.predictions['ml_g1'].flatten() # predictions of g_0(z=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
r0_hat = dml_MLIIVM.predictions['ml_r0'].flatten() # predictions of r_0(z=0, X)
r1_hat = dml_MLIIVM.predictions['ml_r1'].flatten() # predictions of r_0(z=1, X)
r_hat = z * r1_hat + (1 - z) * r0_hat # predictions of r_0
m_hat = dml_MLIIVM.predictions['ml_m'].flatten() # predictions of m_odml_MLIIVM.confint()
```
:::
::::::
\newline














:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# cross-fitted RMSE: outcome
lasso_y_MLIIVM <- sqrt(mean((y-g_hat)^2)) 
lasso_y_MLIIVM

# cross-fitted RMSE: treatment
lasso_d_MLIIVM <- sqrt(mean((d-r_hat)^2)) 
lasso_d_MLIIVM

# cross-fitted RMSE: instrument
lasso_z_MLIIVM <- sqrt(mean((z-m_hat)^2)) 
lasso_z_MLIIVM

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# cross-fitted RMSE: outcome
lasso_y_MLIIVM = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( lasso_y_MLIIVM )

# cross-fitted RMSE: treatment
lasso_d_MLIIVM = np.sqrt( np.mean( ( d - r_hat ) ** 2 ) )
print( lasso_d_MLIIVM )

# cross-fitted RMSE: instrument
lasso_z_MLIIVM = np.sqrt( np.mean( ( z - m_hat ) ** 2 ) ) 
print( lasso_z_MLIIVM )
```
:::
::::::
\newline










Again, we repeat the procedure for the other machine learning methods:


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
### random forest ###

set.seed(123)

lgr::get_logger("mlr3")$set_threshold("warn") 

dml_MLIIVM = DoubleMLIIVM$new(data_IV_aux, 
                              ml_g = randomForest, 
                              ml_m = randomForest_class, 
                              ml_r = randomForest_class, 
                              n_folds=3, 
                              subgroups = list(always_takers = FALSE, 
                                         never_takers = TRUE))


dml_MLIIVM$fit(store_predictions=TRUE)




dml_MLIIVM$summary()

forest_MLIIVM <- dml_MLIIVM$coef
forest_std_MLIIVM <- dml_MLIIVM$se

# predictions

g0_hat <- as.matrix(dml_MLIIVM$predictions$ml_g0) # predictions of g_0(Z=0, X)
g1_hat <- as.matrix(dml_MLIIVM$predictions$ml_g1) # predictions of g_0(Z=1, X)
g_hat <- z*g1_hat+(1-z)*g0_hat # predictions of g_0
r0_hat <- as.matrix(dml_MLIIVM$predictions$ml_r0) # predictions of r_0(Z=0, X)
r1_hat <- as.matrix(dml_MLIIVM$predictions$ml_r1) # predictions of r_0(Z=1, X)
r_hat <- z*r1_hat+(1-z)*r0_hat # predictions of r_0
m_hat <- as.matrix(dml_MLIIVM$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
forest_y_MLIIVM <- sqrt(mean((y-g_hat)^2)) 
forest_y_MLIIVM

# cross-fitted RMSE: treatment
forest_d_MLIIVM <- sqrt(mean((d-r_hat)^2)) 
forest_d_MLIIVM

# cross-fitted RMSE: instrument
forest_z_MLIIVM <- sqrt(mean((z-m_hat)^2)) 
forest_z_MLIIVM


### trees ###

dml_MLIIVM = DoubleMLIIVM$new(data_IV_aux, 
                              ml_g = trees, 
                              ml_m = trees_class, 
                              ml_r = trees_class, 
                              n_folds=3, 
                              subgroups = list(always_takers = FALSE, 
                                         never_takers = TRUE))


dml_MLIIVM$fit(store_predictions=TRUE)




dml_MLIIVM$summary()
tree_MLIIVM <- dml_MLIIVM$coef
tree_std_MLIIVM <- dml_MLIIVM$se

# predictions
g0_hat <- as.matrix(dml_MLIIVM$predictions$ml_g0) # predictions of g_0(Z=0, X)
g1_hat <- as.matrix(dml_MLIIVM$predictions$ml_g1) # predictions of g_0(Z=1, X)
g_hat <- z*g1_hat+(1-z)*g0_hat # predictions of g_0
r0_hat <- as.matrix(dml_MLIIVM$predictions$ml_r0) # predictions of r_0(Z=0, X)
r1_hat <- as.matrix(dml_MLIIVM$predictions$ml_r1) # predictions of r_0(Z=1, X)
r_hat <- z*r1_hat+(1-z)*r0_hat # predictions of r_0
m_hat <- as.matrix(dml_MLIIVM$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
tree_y_MLIIVM <- sqrt(mean((y-g_hat)^2)) 
tree_y_MLIIVM

# cross-fitted RMSE: treatment
tree_d_MLIIVM <- sqrt(mean((d-r_hat)^2)) 
tree_d_MLIIVM

# cross-fitted RMSE: instrument
tree_z_MLIIVM <- sqrt(mean((z-m_hat)^2)) 
tree_z_MLIIVM



### boosting ###

dml_MLIIVM = DoubleMLIIVM$new(data_IV_aux, 
                              ml_g = boost, 
                              ml_m = boost_class, 
                              ml_r = boost_class,
                              n_folds=3, 
                              subgroups = list(always_takers = FALSE, 
                                         never_takers = TRUE))
dml_MLIIVM$fit(store_predictions=TRUE)



dml_MLIIVM$summary()
boost_MLIIVM <- dml_MLIIVM$coef
boost_std_MLIIVM <- dml_MLIIVM$se

# predictions
g0_hat <- as.matrix(dml_MLIIVM$predictions$ml_g0) # predictions of g_0(Z=0, X)
g1_hat <- as.matrix(dml_MLIIVM$predictions$ml_g1) # predictions of g_0(Z=1, X)
g_hat <- z*g1_hat+(1-z)*g0_hat # predictions of g_0
r0_hat <- as.matrix(dml_MLIIVM$predictions$ml_r0) # predictions of r_0(Z=0, X)
r1_hat <- as.matrix(dml_MLIIVM$predictions$ml_r1) # predictions of r_0(Z=1, X)
r_hat <- z*r1_hat+(1-z)*r0_hat # predictions of r_0
m_hat <- as.matrix(dml_MLIIVM$predictions$ml_m) # predictions of m_o

# cross-fitted RMSE: outcome
boost_y_MLIIVM <- sqrt(mean((y-g_hat)^2)) 
boost_y_MLIIVM

# cross-fitted RMSE: treatment
boost_d_MLIIVM <- sqrt(mean((d-r_hat)^2)) 
boost_d_MLIIVM

# cross-fitted RMSE: instrument
boost_z_MLIIVM <- sqrt(mean((z-m_hat)^2)) 
boost_z_MLIIVM

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
### random forest ###
randomForest = RandomForestRegressor(n_estimators=500)
randomForest_class = RandomForestClassifier(n_estimators=500)
np.random.seed(123)
dml_MLIIVM = dml.DoubleMLIIVM(data_IV_aux,
                                  ml_g = randomForest,
                                  ml_m = randomForest_class,
                                  ml_r = randomForest_class,
                                  subgroups = {'always_takers': False,
                                             'never_takers': True},
                                  trimming_threshold = 0.01,
                                  n_folds = 3)

dml_MLIIVM.fit(store_predictions=True)


print( dml_MLIIVM.summary )


forest_MLIIVM = dml_MLIIVM.coef[ 0 ]
forest_std_MLIIVM = dml_MLIIVM.se[ 0 ]

# predictions
g0_hat = dml_MLIIVM.predictions['ml_g0'].flatten() # predictions of g_0(z=0, X)
g1_hat = dml_MLIIVM.predictions['ml_g1'].flatten() # predictions of g_0(z=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
r0_hat = dml_MLIIVM.predictions['ml_r0'].flatten() # predictions of r_0(z=0, X)
r1_hat = dml_MLIIVM.predictions['ml_r1'].flatten() # predictions of r_0(z=1, X)
r_hat = z * r1_hat + (1 - z) * r0_hat # predictions of r_0
m_hat = dml_MLIIVM.predictions['ml_m'].flatten() # predictions of m_o

# cross-fitted RMSE: outcome
forest_y_MLIIVM = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( forest_y_MLIIVM )

# cross-fitted RMSE: treatment
forest_d_MLIIVM = np.sqrt( np.mean( ( d - r_hat ) ** 2 ) )
print( forest_d_MLIIVM )

# cross-fitted RMSE: instrument
forest_z_MLIIVM = np.sqrt( np.mean( ( z - m_hat ) ** 2 ) ) 
print( forest_z_MLIIVM )


### trees ###
np.random.seed(123)
dml_MLIIVM = dml.DoubleMLIIVM(data_IV_aux,
                                  ml_g = trees,
                                  ml_m = trees_class,
                                  ml_r = trees_class,
                                  subgroups = {'always_takers': False,
                                             'never_takers': True},
                                  trimming_threshold = 0.01,
                                  n_folds = 3)
dml_MLIIVM.fit(store_predictions=True)

print( dml_MLIIVM.summary )
tree_MLIIVM = dml_MLIIVM.coef[ 0 ]
tree_std_MLIIVM = dml_MLIIVM.se[ 0 ]


# predictions
g0_hat = dml_MLIIVM.predictions['ml_g0'].flatten() # predictions of g_0(z=0, X)
g1_hat = dml_MLIIVM.predictions['ml_g1'].flatten() # predictions of g_0(z=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
r0_hat = dml_MLIIVM.predictions['ml_r0'].flatten() # predictions of r_0(z=0, X)
r1_hat = dml_MLIIVM.predictions['ml_r1'].flatten() # predictions of r_0(z=1, X)
r_hat = z * r1_hat + (1 - z) * r0_hat # predictions of r_0
m_hat = dml_MLIIVM.predictions['ml_m'].flatten() # predictions of m_o

# cross-fitted RMSE: outcome
tree_y_MLIIVM = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( tree_y_MLIIVM )

# cross-fitted RMSE: treatment
tree_d_MLIIVM = np.sqrt( np.mean( ( d - r_hat ) ** 2 ) )
print( tree_d_MLIIVM )

# cross-fitted RMSE: instrument
tree_z_MLIIVM = np.sqrt( np.mean( ( z - m_hat ) ** 2 ) ) 
print( tree_z_MLIIVM )



### boosting ###
np.random.seed(123)
dml_MLIIVM = dml.DoubleMLIIVM(data_IV_aux,
                                  ml_g = boost,
                                  ml_m = boost_class,
                                  ml_r = boost_class,
                                  subgroups = {'always_takers': False,
                                             'never_takers': True},
                                  trimming_threshold = 0.01,
                                  n_folds = 3)
dml_MLIIVM.fit(store_predictions=True)

print( dml_MLIIVM.summary )
boost_MLIIVM = dml_MLIIVM.coef[ 0 ]
boost_std_MLIIVM = dml_MLIIVM.se[ 0 ]


# predictions
g0_hat = dml_MLIIVM.predictions['ml_g0'].flatten() # predictions of g_0(z=0, X)
g1_hat = dml_MLIIVM.predictions['ml_g1'].flatten() # predictions of g_0(z=1, X)
g_hat = d * g1_hat + ( 1 - d )*g0_hat # predictions of g_0
r0_hat = dml_MLIIVM.predictions['ml_r0'].flatten() # predictions of r_0(z=0, X)
r1_hat = dml_MLIIVM.predictions['ml_r1'].flatten() # predictions of r_0(z=1, X)
r_hat = z * r1_hat + (1 - z) * r0_hat # predictions of r_0
m_hat = dml_MLIIVM.predictions['ml_m'].flatten() # predictions of m_o

# cross-fitted RMSE: outcome
boost_y_MLIIVM = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( boost_y_MLIIVM )

# cross-fitted RMSE: treatment
boost_d_MLIIVM = np.sqrt( np.mean( ( d - r_hat ) ** 2 ) )
print( boost_d_MLIIVM )

# cross-fitted RMSE: instrument
boost_z_MLIIVM = np.sqrt( np.mean( ( z - m_hat ) ** 2 ) ) 
print( boost_z_MLIIVM )
```
:::
::::::
\newline







:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
# cross-fitted RMSE: outcome
lasso_y_MLIIVM <- sqrt(mean((y-g_hat)^2)) 
lasso_y_MLIIVM

# cross-fitted RMSE: treatment
lasso_d_MLIIVM <- sqrt(mean((d-r_hat)^2)) 
lasso_d_MLIIVM

# cross-fitted RMSE: instrument
lasso_z_MLIIVM <- sqrt(mean((z-m_hat)^2)) 
lasso_z_MLIIVM

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# cross-fitted RMSE: outcome
lasso_y_MLIIVM = np.sqrt( np.mean( ( y - g_hat ) ** 2 ) ) 
print( lasso_y_MLIIVM )

# cross-fitted RMSE: treatment
lasso_d_MLIIVM = np.sqrt( np.mean( ( d - r_hat ) ** 2 ) )
print( lasso_d_MLIIVM )

# cross-fitted RMSE: instrument
lasso_z_MLIIVM = np.sqrt( np.mean( ( z - m_hat ) ** 2 ) ) 
print( lasso_z_MLIIVM )
```
:::
::::::
\newline





:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
library(xtable)
table <- matrix(0, 5, 4)
table[1,1:4]   <- c(lasso_MLIIVM,forest_MLIIVM,tree_MLIIVM,boost_MLIIVM)
table[2,1:4]   <- c(lasso_std_MLIIVM,forest_std_MLIIVM,tree_std_MLIIVM,boost_std_MLIIVM)
table[3,1:4]   <- c(lasso_y_MLIIVM,forest_y_MLIIVM,tree_y_MLIIVM,boost_y_MLIIVM)
table[4,1:4]   <- c(lasso_d_MLIIVM,forest_d_MLIIVM,tree_d_MLIIVM,boost_d_MLIIVM)
table[5,1:4]   <- c(lasso_z_MLIIVM,forest_z_MLIIVM,tree_z_MLIIVM,boost_z_MLIIVM)
rownames(table) <- c("Estimate","Std.Error","RMSE Y","RMSE D","RMSE Z")
colnames(table) <- c("Lasso","Random Forest","Trees","Boosting")
tab<- xtable(table, digits = 2)
tab
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
table3 = np.zeros( (5, 4) )
table3[0,0:4] = lasso_MLIIVM,forest_MLIIVM,tree_MLIIVM,boost_MLIIVM
table3[1,0:4] = lasso_std_MLIIVM,forest_std_MLIIVM,tree_std_MLIIVM,boost_std_MLIIVM
table3[2,0:4] = lasso_y_MLIIVM,forest_y_MLIIVM,tree_y_MLIIVM,boost_y_MLIIVM
table3[3,0:4] = lasso_d_MLIIVM,forest_d_MLIIVM,tree_d_MLIIVM,boost_d_MLIIVM
table3[4,0:4] = lasso_z_MLIIVM,forest_z_MLIIVM,tree_z_MLIIVM,boost_z_MLIIVM
table3_pd = pd.DataFrame( table3 , 
			      index = ["Estimate","Std.Error","RMSE Y","RMSE D","RMSE Z" ], \
		        columns = [ "Lasso","Random Forest","Trees","Boosting"])

table3_pd.to_latex()
```
:::
::::::
\newline


We report results based on four ML methods for estimating the nuisance functions used in
forming the orthogonal estimating equations. We find again that the estimates of the treatment effect are stable across ML methods. The estimates are highly significant, hence we would reject the hypothesis
that the effect of 401(k) participation has no effect on financial health.

We might rerun the model using the best ML method for each equation to get a final estimate for the treatment effect of participation:


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}"}
```{r  message=FALSE, warning=FALSE}
set.seed(123)
lgr::get_logger("mlr3")$set_threshold("warn") 
dml_MLIIVM = DoubleMLIIVM$new(data_IV_aux, 
                              ml_g = randomForest, 
                              ml_m = lasso_class, 
                              ml_r = lasso_class,
                              n_folds=3, 
                              subgroups = list(always_takers = FALSE, 
                                         never_takers = TRUE))



dml_MLIIVM$fit(store_predictions=TRUE)



dml_MLIIVM$summary()
best_MLIIVM <- dml_MLIIVM$coef
best_std_MLIIVM <- dml_MLIIVM$se
```

:::
::: {.column width="1%" datI a-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python}
# Random best
np.random.seed(123)
dml_MLIIVM = dml.DoubleMLIIVM(data_IV_aux,
                                  ml_g = randomForest,
                                  ml_m = lasso_class,
                                  ml_r = lasso_class,
                                  subgroups = {'always_takers': False,
                                             'never_takers': True},
                                  trimming_threshold = 0.01,
                                  n_folds = 3)
dml_MLIIVM.fit(store_predictions=True)

print( dml_MLIIVM.summary )

best_MLIIVM = dml_MLIIVM.coef[ 0 ]
best_std_MLIIVM = dml_MLIIVM.se[ 0 ]
```
:::
::::::
\newline






