# Using Dagitty in the Analysis of Impact of 401(k) on Net Financial Wealth

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>




:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE, fig.height=5, fig.width=15}
# install and load package
# install.packages("dagitty")
# install.packages("ggdag")
library(dagitty)
library(ggdag)

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline



## Graphs for 401(K) Analsyis


Here we have <br />

* $Y$ -- net financial assets;
* $X$ -- worker characteristics (income, family size, other retirement plans; see lecture notes for details);
* $F$ -- latent (unobserved) firm characteristics
* $D$ -- 401(K) eligibility, deterimined by $F$ and $X$



## State one graph (where F determines X) and plot it

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
#generate a DAGs and plot them

G1 = dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [uobserved, pos="0, -1"]
D -> Y
X -> D
F -> X
F -> D
X -> Y}')


ggdag(G1)+  theme_dag()
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline


## List minimal adjustment sets to identify causal effecs $D \to Y$

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
adjustmentSets( G1, "D", "Y",effect="total" ) 
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline


## What is the underlying principle? 

Here condition on X blocks backdoor paths from Y to D (Pearl).  Dagitty correctly finds X (and does many more correct decisions, when we consider more elaborate structures. Why do we want to consider more elaborate structures? The very empirical problem requires us to do so!

## Another Graph (wherere $X$ determine $F$):

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
#generate a couple of DAGs and plot them

G2 = dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [uobserved, pos="0, -1"]
D -> Y
X -> D
X -> F
F -> D
X -> Y}')


ggdag(G2)+  theme_dag()
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline









:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
adjustmentSets( G2, "D", "Y", effect="total" )
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline








## One more graph (encompassing previous ones), where (F, X) are jointly determined by latent factors $A$. We can allow in fact the whole triple (D, F, X) to be jointly determined by latent factors $A$.


This is much more realistic graph to consider.


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
G3 = dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
D -> Y
X -> D
F -> D
A -> F
A -> X
A -> D
X -> Y}')

adjustmentSets( G3, "D", "Y", effect="total"  ) 

ggdag(G3)+  theme_dag()
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline




## Threat to Idenitification: What if $F$ also directly affects $Y$? (Note that there are no valid adjustment sets in this case)


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
G4 = dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
D -> Y
X -> D
F -> D
A -> F
A -> X
A -> D
F -> Y
X -> Y}')


ggdag(G4)+  theme_dag()
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
adjustmentSets( G4, "D", "Y",effect="total"  )

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline


Note that no output means that there is no valid adustment set (among observed variables)

<span style="font-size:2em; font-weight:bold">  How can F affect Y directly? Is it reasonable? </span>

## Introduce Match Amount $M$ (very important mediator, why mediator?). $M$ is not observed.  Luckily adjusting for $X$ still works if there is no $F \to M$ arrow.

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
G5 = dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
M [unobserved, pos="2, -.5"]
D -> Y
X -> D
F -> D
A -> F
A -> X
A -> D
D -> M
M -> Y
X -> M
X -> Y}')

print( adjustmentSets( G5, "D", "Y",effect="total"  ) )

ggdag(G5)+  theme_dag()
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline


## If  there is $F \to M$ arrow, then adjusting for $X$ is not sufficient.


:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
G6 = dagitty('dag{
Y [outcome,pos="4, 0"]
D [exposure,pos="0, 0"]
X [confounder, pos="2,-2"]
F [unobserved, pos="0, -1"]
A [unobserved, pos="-1, -1"]
M [uobserved, pos="2, -.5"]
D -> Y
X -> D
F -> D
A -> F
A -> X
D -> M
F -> M
A -> D
M -> Y
X -> M
X -> Y}')

print( adjustmentSets( G6, "D", "Y" ),effect="total"  )

ggdag(G6)+  theme_dag()
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

```
:::
::::::
\newline

## Question:
 
Given the analysis above, do you find the adjustment for workers' characteristics a credible strategy to identify the causal (total effect) of 401 (k) elligibility on net financial wealth?

     * If yes, click an "upvote" button at the top
     * If no, please click an "upvote" button at the top