# A Simple Example of Properties of IV estimator when Instruments are Weak

Simulation Design

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>




:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE, fig.height=5, fig.width=15}
# Simulation Design

library(hdm)
set.seed(1)
B= 10000 # trials
IVEst = rep(0, B)
n=100
beta = .25   # .2 weak IV
#beta = 1   #   1 strong IV


U =  rnorm(n)  
Z = rnorm(n)  #generate instrument
D = beta*Z + U  #generate endogenougs variable
Y =  D+ U  # the true causal effect is 1


summary(lm(D~Z))  # first stage is very weak here
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}
import hdmpy
import numpy as np
import random
import statsmodels.api as sm
import matplotlib.pyplot as plt
import numpy as np
from matplotlib import colors
from statsmodels.sandbox.regression.gmm import IV2SLS
import numpy as np                                                              
import seaborn as sns                                                           
from scipy import stats                                                         
import matplotlib.pyplot as plt
import warnings
warnings.filterwarnings("ignore")

# Simulation Design

# Set seed
np.random.seed(0)
B = 1000
IVEst = np.zeros( B )
n = 100
beta = .25

mean = 0
sd = 1

U = np.random.normal( mean , sd, n ).reshape( n, 1 )
Z = np.random.normal( mean , sd, n ).reshape( n, 1 )
D = beta*Z + U 
Y = D + U

mod = sm.OLS(D, sm.add_constant(Z))    # Describe model
res = mod.fit()
print(res.summary())

```
:::
::::::
\newline










:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
summary(tsls(x=NULL, d=D, y=Y, z=Z))  #
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}
IV = IV2SLS(Y, D, sm.add_constant(Z))
IV_res = IV.fit()
print(IV_res.summary())
```
:::
::::::
\newline


Note that the instrument is weak here (contolled by $\beta$) -- the t-stat is less than 4.

## Run 1000 trials to evaluate distribution of the IV estimator

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}

# Simulation Design

set.seed(1)
B= 10000 # trials
IVEst = rep(0, B)

for(i in 1:B){
U =  rnorm(n)  
Z = rnorm(n)  #generate instrument
D = beta*Z + U  #generate endogenougs variable
Y =  D+ U  # the true causal effect is 1
IVEst[i] = coef(tsls(x=NULL, d=D, y=Y, z=Z))[1,1]
}



```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

# Simulation design 

# Set seed
np.random.seed(0)
B = 1000 # Trials
IVEst = np.zeros( B )

for i in range( 0, B ):
    U = np.random.normal( mean , sd, n ).reshape( n, 1 )
    Z = np.random.normal( mean , sd, n ).reshape( n, 1 )
    D = beta*Z + U 
    Y = D + U
    
    IV = IV2SLS(Y, D, sm.add_constant(Z))
    IV_res = IV.fit()
    
    IVEst[ i ] = IV_res.summary2().tables[1]["Coef."][0]
    
```
:::
::::::
\newline





## Plot the Actual Distribution against the Normal Approximation (based on Strong Instrument Assumption)



:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
plot(density(IVEst-1, n=1000, from=-5, to=5),col=4, xlim= c(-5, 5),  
     xlab= "IV Estimator -True Effect", main="Actual Distribution vs Gaussian")

val=seq(-5, 5, by=.05)
var = (1/beta^2)*(1/100) # theoretical variance of IV
sd = sqrt(var)
lines(val, dnorm(val, sd=sd), col=2, lty=2)

rejection.frequency = sum(( abs(IVEst-1)/sd > 1.96))/B

```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}

val = np.arange(-5,5.5,0.05)
var = (1/beta**2)*(1/100)   # theoretical variance of IV
sd = np.sqrt(var)

normal_dist = np.random.normal(0,sd,val.shape[0])

# plotting both distibutions on the same figure
fig = sns.kdeplot(IVEst-1, shade=True, color="r")
fig = sns.kdeplot(normal_dist, shade=True, color="b")

plt.title("Actual Distribution vs Gaussian")
plt.xlabel('IV Estimator -True Effect')
plt.xlim(-5,5)
    
```
:::
::::::
\newline





:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
cat(c("Rejection Frequency is ", 
      rejection.frequency, " while we expect it to be .05"))
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}
rejection_frequency = np.sum(( np.abs(IVEst-1)/sd > 1.96))/B
print( f"Rejection Frequency is {rejection_frequency} ,while we expect it to be .05")
  
```
:::
::::::
\newline



## Some Help Functions

:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
help(tsls)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}
help(IV2SLS)
```
:::
::::::
\newline



:::::: {.columns}
::: {.column width="49.5%" data-latex="{0.48\textwidth}" }
```{r  message=FALSE, warning=FALSE}
help(density)
```

:::
::: {.column width="1%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="49.5%" data-latex="{0.48\textwidth}"}

```{python fig.height=5, fig.width=15}
help(sns.kdeplot)
```
:::
::::::
\newline
