--- 
title: "Machine Learning and Causal Inference"
author: "Alexander Quispe"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---

# Predictive Inference 1

In labor economics an important question is what determines the wage of workers. This is a causal question, but we could begin to investigate from a predictive perspective.

In the following wage example, $Y$ is the hourly wage of a worker and $X$ is a vector of worker's characteristics, e.g., education, experience, gender.
Two main questions here are:    

* How to use job-relevant characteristics, such as education and experience, to best predict wages?

* What is the difference in predicted wages between men and women with the same job-relevant characteristics?

In this lab, we focus on the prediction question first.

## Data

The data set we consider is from the March Supplement of the U.S. Current Population Survey, year 2015. We select white non-hispanic individuals, aged 25 to 64 years, and working more than 35 hours per week during at least 50 weeks of the year. We exclude self-employed workers; individuals living in group quarters; individuals in the military, agricultural or private household sectors; individuals with inconsistent reports on earnings and employment status; individuals with allocated or missing information in any of the variables used in the analysis; and individuals with hourly wage below $3$.

The variable of interest $Y$ is the hourly wage rate constructed as the ratio of the annual earnings to the total number of hours worked, which is constructed in turn as the product of number of weeks worked and the usual number of hours worked per week. In our analysis, we also focus on single (never married) workers. The final sample is of size $n = 5150$.

## Data Analysis

### R code

1. Import relevant packages

There are not additional packages to import.

```{r packages,echo= TRUE,message=FALSE,warning=FALSE}
library(dplyr)
library(kableExtra)
```

2. We start loading the data set.

```{r import-data,echo= TRUE}
load("./data/wage2015_subsample_inference.Rdata")
```

The dimensions are `r dim(data)[1]` rows and `r dim(data)[2]` columns.

Let's have a look at the structure of the data.

```{r type-table, echo=TRUE}

# Calculate the means and convert it into a dataframe
table0 <- data.frame(apply(data,2,class))%>%
  rename(Type =`apply.data..2..class.`)

table0$Variables <- rownames(table0)
rownames(table0) <- c(1:nrow(table0))

# Table presentation
table0 %>% select(2,1)%>%
  kable("markdown",caption = "Type of the Variables")
```

3. Give structure to the variables

We are constructing the output variable $Y$ and the matrix $Z$ which includes the characteristics of workers that are given in the data.

```{r variables-structure, echo=TRUE}

# Calculate the log wage.
Y <- log(data$wage)
# Number of observaciones
n <- length(Y)
# Regressors
Z <- data[,c("wage","lwage")]
# Number of regressors
p <- dim(Z)[2]
```

Number of observation: `r n`
Number of raw regressors:`r p`

For the outcome variable wage and a subset of the raw regressors, we calculate the empirical mean to get familiar with the data.

```{r table-one, echo = TRUE}
# Select the variables.
Z_subset <- data[,c("lwage","sex","shs","hsg","scl","clg","ad","mw","so","we","ne","exp1")]

# Create table
table1 <- data.frame(as.numeric(lapply(Z_subset,mean))) %>%
  mutate(Variables = c("Log Wage","Sex","Some High School","High School Graduate","Some College","College Graduate", "Advanced Degree","Midwest","South","West","Northeast","Experience")) %>%
  rename(`Sample Mean` = `as.numeric.lapply.Z_subset..mean..`) %>%
  select(2,1)

# HTML table
table1 %>%
  kable("markdown",caption = "Descriptive Statistics")
```

E.g., the share of female workers in our sample is $~44% (sex=1$ if female).

Alternatively, we can also print the table as latex.

```{r table-one-latex, echo=TRUE}
print(table1, type="latex")
```

## Prediction Question

Now, we will construct a prediction rule for hourly wage $Y$ , which depends linearly on job-relevant characteristics  $X$:

$$Y = \beta ′X + \epsilon $$
 
Our goals are

* Predict wages using various characteristics of workers.

* Assess the predictive performance using the (adjusted) sample MSE, the (adjusted) sample $R^2$ and the out-of-sample $MSE$ and $R^2$.

We employ two different specifications for prediction:

- **Basic Model**: $X$ consists of a set of raw regressors (e.g. gender, experience, education indicators, occupation and industry indicators, regional indicators).

- **Flexible Model**: $X$ consists of all raw regressors from the basic model plus occupation and industry indicators, transformations (e.g.,$exp2$ and $exp3$) and additional two-way interactions of polynomial in experience with other regressors. An example of a regressor created through a two-way interaction is experience times the indicator of having a college degree.

Using the **Flexible Model**, enables us to approximate the real relationship by a more complex regression model and therefore to reduce the bias. The **Flexible Model** increases the range of potential shapes of the estimated regression function. In general, flexible models often deliver good prediction accuracy but give models which are harder to interpret.

Now, let us fit both models to our data by running ordinary least squares (ols):

```{r basic-model}
basic <- lwage~ (sex + exp1 + shs + hsg+ scl + clg + mw + so + we+occ2+ind2)
regbasic <- lm(basic, data=data)
regbasic # estimated coefficients
```

Number of regressors in the basic model **`r length(regbasic$coef)`**.

```{r flexible-model}
flex <- lwage ~ sex + shs+hsg+scl+clg+occ2+ind2+mw+so+we + (exp1+exp2+exp3+exp4)*(shs+hsg+scl+clg+occ2+ind2+mw+so+we)
regflex <- lm(flex, data=data)
regflex # estimated coefficients
```

Number of regressors in the flexible model:**`r length(regflex$coef)`**.

- **Lasso Model**

First, we import a essential library.

```{r hdm-packages, message=FALSE, warning=FALSE}
library(hdm)
```


```{r}
flex <- lwage ~ sex + shs+hsg+scl+clg+occ2+ind2+mw+so+we + (exp1+exp2+exp3+exp4)*(shs+hsg+scl+clg+occ2+ind2+mw+so+we)
lassoreg<- rlasso(flex, data=data)

sumlasso<- summary(lassoreg)
```

Now, we can evaluate the performance of both models based on the (adjusted) $R^2_{sample}$  and the (adjusted) $MSE_{sample}$:





