--- 
title: "Machine Learning and Causal Inference"
author: "Alexander Quispe"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---

# Predictive Inference 1

In labor economics an important question is what determines the wage of workers. This is a causal question, but we could begin to investigate from a predictive perspective.

In the following wage example, $Y$ is the hourly wage of a worker and $X$ is a vector of worker's characteristics, e.g., education, experience, gender.
Two main questions here are:    

* How to use job-relevant characteristics, such as education and experience, to best predict wages?

* What is the difference in predicted wages between men and women with the same job-relevant characteristics?

In this lab, we focus on the prediction question first.

## Data

The data set we consider is from the March Supplement of the U.S. Current Population Survey, year 2015. We select white non-hispanic individuals, aged 25 to 64 years, and working more than 35 hours per week during at least 50 weeks of the year. We exclude self-employed workers; individuals living in group quarters; individuals in the military, agricultural or private household sectors; individuals with inconsistent reports on earnings and employment status; individuals with allocated or missing information in any of the variables used in the analysis; and individuals with hourly wage below $3$.

The variable of interest $Y$ is the hourly wage rate constructed as the ratio of the annual earnings to the total number of hours worked, which is constructed in turn as the product of number of weeks worked and the usual number of hours worked per week. In our analysis, we also focus on single (never married) workers. The final sample is of size $n = 5150$.

## Data Analysis

### R and Python code

1. Import relevant packages

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**
```{r packages,echo= TRUE,message=FALSE,warning=FALSE}
library(dplyr)
library(kableExtra)
```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

**Python code**
```{python python-packages}
# import pandas as pd
# import numpy as np
# import pyreadr
```
:::
::::::

2. We start loading the data set.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
**R code**

```{r import-data,echo= TRUE}
load("./data/wage2015_subsample_inference.Rdata")
```
:::

::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**

```{python import-python-data}
#rdata_read = pyreadr.read_r("../data/wage2015_subsample_inference.Rdata")
#data = rdata_read[ 'data' ]
#type(data)
#data.shape
```
:::
::::::

The dimensions are `r dim(data)[1]` rows and `r dim(data)[2]` columns.

Let's have a look at the structure of the data.

**R code**
```{r type-table, echo=TRUE}

# Calculate the means and convert it into a dataframe
table0 <- data.frame(lapply(data,class))%>%
  tidyr::gather("Variable","Type")

# Table presentation
table0 %>%
  kable("markdown",caption = "Type of the Variables")
```

**Python code**

```{python type-table-python}
# data.info()

# data.describe()
```


3. Give structure to the variables

We are constructing the output variable $Y$ and the matrix $Z$ which includes the characteristics of workers that are given in the data.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
**R code**

```{r variables-structure, echo=TRUE}

# Calculate the log wage.
Y <- log(data$wage)
# Number of observaciones
n <- length(Y)
# Regressors
Z <- data[,c("wage","lwage")]
# Number of regressors
p <- dim(Z)[2]
```

Number of observation: `r n`  
Number of raw regressors:`r p`
:::

::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**

```{python variables-structure-python}
## Calculate the log wage.
# Y = np.log2(data['wage']) 
## Number of observaciones
# n = len(Y)
## Regressors
# z = data.loc[:, ~data.columns.isin(['wage', 'lwage','Unnamed: 0'])]
## Number of regressors
# p = z.shape[1]
# print("Number of observation:", n, '\n')
# print( "Number of raw regressors:", p)
```
:::
::::::


4. For the outcome variable wage and a subset of the raw regressors, we calculate the empirical mean to get familiar with the data.


**R code**
```{r table-one, echo = TRUE}
# Select the variables.
Z_subset <- data[,c("lwage","sex","shs","hsg","scl","clg","ad","mw","so","we","ne","exp1")]
# Create table
table1 <- data.frame(as.numeric(lapply(Z_subset,mean))) %>%
  mutate(Variables = c("Log Wage","Sex","Some High School","High School Graduate","Some College","College Graduate", "Advanced Degree","Midwest","South","West","Northeast","Experience")) %>%
  rename(`Sample Mean` = `as.numeric.lapply.Z_subset..mean..`) %>%
  select(2,1)
# HTML table
table1 %>%
  kable("markdown",caption = "Descriptive Statistics")
```

**Python code**

```{python table-one-python}
# Z_subset = data.loc[:, data.columns.isin(["lwage","sex","shs","hsg","scl","clg","ad","mw","so","we","ne","exp1"])]
# table = Z_subset.mean(axis=0)
# table
```


```{python table-one-final-python}
# table = pd.DataFrame(data=table, columns={"Sample mean":"0"} )
# table.index
# index1 = list(table.index)
# index2 = ["Log Wage","Sex","Some High School","High School Graduate",\
#           "Some College","College Graduate", "Advanced Degree","Midwest",\
#           "South","West","Northeast","Experience"]

# table = table.rename(index=dict(zip(index1,index2)))
# table
```

E.g., the share of female workers in our sample is ~44% ($sex=1$ if female).

Alternatively, we can also print the table as latex.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
**R code**

```{r table-one-latex, echo=TRUE}
print(table1, type="latex")
```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**
```{python table-one-latex-python}
# print(table.to_latex())
```

:::
::::::

## Prediction Question

Now, we will construct a prediction rule for hourly wage $Y$ , which depends linearly on job-relevant characteristics  $X$:

$$Y = \beta ′X + \epsilon $$
 
Our goals are

* Predict wages using various characteristics of workers.

* Assess the predictive performance using the (adjusted) sample MSE, the (adjusted) sample $R^2$ and the out-of-sample $MSE$ and $R^2$.

We employ two different specifications for prediction:

- **Basic Model**: $X$ consists of a set of raw regressors (e.g. gender, experience, education indicators, occupation and industry indicators, regional indicators).

- **Flexible Model**: $X$ consists of all raw regressors from the basic model plus occupation and industry indicators, transformations (e.g.,$exp2$ and $exp3$) and additional two-way interactions of polynomial in experience with other regressors. An example of a regressor created through a two-way interaction is experience times the indicator of having a college degree.

Using the **Flexible Model**, enables us to approximate the real relationship by a more complex regression model and therefore to reduce the bias. The **Flexible Model** increases the range of potential shapes of the estimated regression function. In general, flexible models often deliver good prediction accuracy but give models which are harder to interpret.

Now, let us fit both models to our data by running ordinary least squares (ols):


### Basic Model:

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**

```{r basic-model}
basic <- lwage~ (sex + exp1 + shs + hsg+ scl + clg + mw + so + we+occ2+ind2)
regbasic <- lm(basic, data=data)
regbasic # estimated coefficients
```

Number of regressors in the basic model **`r length(regbasic$coef)`**.
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

**Python**

Regression packages:
```{python ols-packages-python}
# import statsmodels.api as sm
# import statsmodels.formula.api as smf
```

```{python basic-model-python}
# basic = 'lwage ~ sex + exp1 + shs + hsg+ scl + clg + mw + so + we + occ2+ ind2'
# basic_results = smf.ols(basic , data=data).fit()
# print(basic_results.summary()) # estimated coefficients
```

```{python basic-model-python-regressors}
# print( "Number of regressors in the basic model:",len(basic_results.params), '\n')  # number of regressors in the Basic Model
```

:::
::::::

### Flexible Model:

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**
```{r flexible-model}
flex <- lwage ~ sex + shs+hsg+scl+clg+occ2+ind2+mw+so+we + (exp1+exp2+exp3+exp4)*(shs+hsg+scl+clg+occ2+ind2+mw+so+we)
regflex <- lm(flex, data=data)
regflex # estimated coefficients
```

Number of regressors in the flexible model:**`r length(regflex$coef)`**.
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

**Python code**
```{python flexible-model-python}
# flex = 'lwage ~ sex + shs+hsg+scl+clg+occ2+ind2+mw+so+we + (exp1+exp2+exp3+exp4)*(shs+hsg+scl+clg+occ2+ind2+mw+so+we)'
# flex_results_0 = smf.ols(flex , data=data)
# flex_results = smf.ols(flex , data=data).fit()
# print(flex_results.summary()) # estimated coefficients
# print( "Number of regressors in the basic model:",len(flex_results.params), '\n')
```
:::
::::::

### Lasso Model:

First, we import the essential libraries.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
**R code**

```{r hdm-packages, message=FALSE, warning=FALSE}
library(hdm)
```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**

```{python lasso-packages-python}
# from sklearn.linear_model import LassoCV
# from sklearn import linear_model
# from sklearn.preprocessing import PolynomialFeatures
# from sklearn.metrics import mean_squared_error
```

:::
::::::

Then, we estimate the lasso model.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**
```{r lasso-estimation}
flex <- lwage ~ sex + shs+hsg+scl+clg+occ2+ind2+mw+so+we + (exp1+exp2+exp3+exp4)*(shs+hsg+scl+clg+occ2+ind2+mw+so+we)
lassoreg<- rlasso(flex, data=data)

sumlasso<- summary(lassoreg)
```

:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**

```{python exogenous-var-python}
# Get exogenous variables from flexible model
# X = flex_results_0.exog
# X.shape
# 
# # Set endogenous variable
# lwage = data["lwage"]
# lwage.shape
```

```{python lasso-estimation-python}
# alpha=0.1
# # Set penalty value = 0.1
# #reg = linear_model.Lasso(alpha=0.1/np.log(len(lwage)))
# reg = linear_model.Lasso(alpha = alpha)
# 
# # LASSO regression for flexible model
# reg.fit(X, lwage)
# lwage_lasso_fitted = reg.fit(X, lwage).predict( X )
# 
# # coefficients 
# reg.coef_
# print('Lasso Regression: R^2 score', reg.score(X, lwage))
```

:::
::::::

Now, we can evaluate the performance of both models based on the (adjusted) $R^2_{sample}$  and the (adjusted) $MSE_{sample}$:

1. R-Squared $(R^2)$

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
**R code**
```{r r-squared}
# Summary from basic and flexible model.
sumbasic <- summary(regbasic)
sumflex <- summary(regflex)

#  R-squared from basic, flexible and lasso models
R2.1 <- sumbasic$r.squared
R2.adj1 <- sumbasic$adj.r.squared

R2.2 <- sumflex$r.squared
R2.adj2 <- sumflex$adj.r.squared

R2.L <- sumlasso$r.squared
R2.adjL <- sumlasso$adj.r.squared
```

- R-squared for the basic model: `r R2.1`.
- Adjusted R-squared for the basic model: `r R2.adj1`.
- R-squared for the flexible model: `r R2.2`.
- Adjusted R-squared for the flexible model: `r R2.adj2`.
- R-squared for the lasso with flexible model: `r R2.L`.
- Adjusted R-squared for the flexible model: `r R2.adjL`.

:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

**Python**

```{python r-squared-python-two}
# Assess the predictive performance
# R2_1 = basic_results.rsquared
# print("R-squared for the basic model: ", R2_1, "\n")
# R2_adj1 = basic_results.rsquared_adj
# print("adjusted R-squared for the basic model: ", R2_adj1, "\n")
# 
# R2_2 = flex_results.rsquared
# print("R-squared for the basic model: ", R2_2, "\n")
# R2_adj2 = flex_results.rsquared_adj
# print("adjusted R-squared for the basic model: ", R2_adj2, "\n")
# R2_L = reg.score(flex_results_0.exog, lwage)
# print("R-squared for LASSO: ", R2_L, "\n")
# R2_adjL = 1 - (1-R2_L)*(len(lwage)-1)/(len(lwage)-X.shape[1]-1)
# print("adjusted R-squared for LASSO: ", R2_adjL, "\n")
```

:::
::::::

2. Mean Squared Error $MSE$

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
**R code**
```{r mean-squared-error}
MSE1 <- mean(sumbasic$res^2)
p1 <- sumbasic$df[1] # number of regressors
MSE.adj1 <- (n/(n-p1))*MSE1


MSE2 <-mean(sumflex$res^2)
p2 <- sumflex$df[1]
MSE.adj2 <- (n/(n-p2))*MSE2

MSEL <-mean(sumlasso$res^2)
pL <- length(sumlasso$coef)
MSE.adjL <- (n/(n-pL))*MSEL

```

* MSE for the basic model: `r MSE1`
* Adjusted MSE for the basic model: `r MSE.adj1`
* MSE for the flexible model: `r MSE2`
* Adjusted MSE for the flexible model: `r MSE.adj2` 
* MSE for the lasso flexible model: `r MSEL`
* Adjusted MSE for the lasso flexible model: `r MSE.adjL`


:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**

```{python mean-squared-error-python}
# calculating the MSE
# MSE1 =  np.mean(basic_results.resid**2)
# print("MSE for the basic model: ", MSE1, "\n")
# p1 = len(basic_results.params) # number of regressors
# n = len(lwage)
# MSE_adj1  = (n/(n-p1))*MSE1
# print("adjusted MSE for the basic model: ", MSE_adj1, "\n")
# 
# MSE2 =  np.mean(flex_results.resid**2)
# print("MSE for the flexible model: ", MSE2, "\n")
# p2 = len(flex_results.params) # number of regressors
# n = len(lwage)
# MSE_adj2  = (n/(n-p2))*MSE2
# print("adjusted MSE for the flexible model: ", MSE_adj2, "\n")
# 
# 
# MSEL = mean_squared_error(lwage, lwage_lasso_fitted)
# print("MSE for the LASSO model: ", MSEL, "\n")
# pL = reg.coef_.shape[0] # number of regressors
# n = len(lwage)
# MSE_adjL  = (n/(n-pL))*MSEL
# print("adjusted MSE for LASSO model: ", MSE_adjL, "\n")
```
:::
::::::

Latex presentation

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**
```{r table-two}
Models <- c("Basic reg","Flexible reg","Lasso reg")
p <- c(p1,p2,pL)
R_2 <- c(R2.1,R2.2,R2.L)
MSE <- c(MSE1,MSE2,MSEL)
R_2_adj <- c(R2.adj1,R2.adj2,R2.adjL)
MSE_adj <- c(MSE.adj1,MSE.adj2,MSE.adjL)
```

```{r table-two-html}
data.frame(Models,p,R_2,MSE,R_2_adj,MSE_adj) %>%
  kable("markdown",caption = "Descriptive Statistics")

```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**
```{python table-two-python}
# import array_to_latex as a2l
# 
# table = np.zeros((3, 5))
# table[0,0:5] = [p1, R2_1, MSE1, R2_adj1, MSE_adj1]
# table[1,0:5] = [p2, R2_2, MSE2, R2_adj2, MSE_adj2]
# table[2,0:5] = [pL, R2_L, MSEL, R2_adjL, MSE_adjL]
# table
```

```{python table-two-latex}
# table = pd.DataFrame(table, columns = ["p","$R^2_{sample}$","$MSE_{sample}$","$R^2_{adjusted}$", "$MSE_{adjusted}$"], index = ["basic reg","flexible reg", "lasso flex"])
# table
```
:::
::::::

Considering all measures above, the flexible model performs slightly better than the basic model.

One procedure to circumvent this issue is to use data splitting that is described and applied in the following.

## Data Splitting

Measure the prediction quality of the two models via data splitting:

- Randomly split the data into one training sample and one testing sample. Here we just use a simple method (stratified splitting is a more sophisticated version of splitting that we can consider).

- Use the training sample for estimating the parameters of the **Basic Model** and the **Flexible Model**.

- Use the testing sample for evaluation. Predict the $wage$ of every observation in the testing sample based on the estimated parameters in the training sample.

- Calculate the **Mean Squared Prediction Error** $MSE_{test}$ based on the testing sample for both prediction models.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**
```{r data-spliting}
# to make the results replicable (generating random numbers)
set.seed(1)
# draw (4/5)*n random numbers from 1 to n without replacing them
random_2 <- sample(1:n, floor(n*4/5))
# training sample
train <- data[random_2,]
# testing sample
test <- data[-random_2,]
```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
**Python code**
```{python data-splitting-python}
# # Import relevant packages for splitting data
# import random
# import math
# 
# # Set Seed
# # to make the results replicable (generating random numbers)
# np.random.seed(0)
# random = np.random.randint(0,n, size=math.floor(n))
# data["random"] = random
# random # the array does not change 
```

```{python random-data-splitting-python}
# data_2 = data.sort_values(by=['random'])
# data_2.head()
```

```{python training-testing-python}
# Create training and testing sample 
# train = data_2[ : math.floor(n*4/5)]    # training sample
# test =  data_2[ math.floor(n*4/5) : ]   # testing sample
# print(train.shape)
# print(test.shape)
```
:::
::::::

The train data dimensions are `r dim(train)[1]` rows and `r dim(train)[2]` columns.

We estimate the parameters using the training data set.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
```{r basic-training-estimate}
# basic model
# estimating the parameters in the training sample
regbasic <- lm(basic, data=train)
regbasic
```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

```{python}
# # basic model
# # estimating the parameters in the training sample
# basic_results = smf.ols(basic , data=train).fit()
# print(basic_results.summary())
```
:::
::::::

Then predict using the parameters in the testing sample. 

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}
```{r basic-testing-predict}
# calculating the out-of-sample MSE
trainregbasic <- predict(regbasic, newdata=test)
trainregbasic
```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}
```{python}
# lwage_test = test["lwage"].values
# test = sm.add_constant(test)   #add constant 
# 
# lwage_pred =  basic_results.predict(test) # predict out of sample
# print(lwage_pred)
```

:::
::::::

Finally, we test the predictions.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**
```{r basic-mse}
y.test <- log(test$wage)
MSE.test1 <- sum((y.test-trainregbasic)^2)/length(y.test)
R2.test1<- 1- MSE.test1/var(y.test)
```

* Test MSE for the basic model: `r MSE.test1`.
* Test R2 for the basic model: `r R2.test1`.

:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

**Python code**
```{python}
# MSE_test1 = np.sum((lwage_test-lwage_pred)**2)/len(lwage_test)
# R2_test1  = 1 - MSE_test1/np.var(lwage_test)
# 
# print("Test MSE for the basic model: ", MSE_test1, " ")
# print("Test R2 for the basic model: ", R2_test1)
```

:::
::::::

In the basic model, the $MSE test$ is quite closed to the $MSE sample$.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

```{r fexible-estimate-prediction}

# estimating the parameters
regflex <- lm(flex, data=train)

# calculating the out-of-sample MSE
trainregflex<- predict(regflex, newdata=test)

y.test <- log(test$wage)
MSE.test2 <- sum((y.test-trainregflex)^2)/length(y.test)
R2.test2<- 1- MSE.test2/var(y.test)
```

* Test MSE for the flexible model: `r MSE.test2`.

* Test R2 for the flexible model: `r R2.test2`.

:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

```{python}
# # Flexible model
# # estimating the parameters in the training sample
# flex_results = smf.ols(flex , data=train).fit()
# 
# # calculating the out-of-sample MSE
# lwage_flex_pred =  flex_results.predict(test) # predict out of sample
# lwage_test = test["lwage"].values
# 
# MSE_test2 = np.sum((lwage_test-lwage_flex_pred)**2)/len(lwage_test)
# R2_test2  = 1 - MSE_test2/np.var(lwage_test)
# 
# print("Test MSE for the flexible model: ", MSE_test2, " ")
# print("Test R2 for the flexible model: ", R2_test2)
```

:::
::::::

In the flexible model, the discrepancy between the $MSE_{test}$ and the $MSE_{sample}$ is not large.

It is worth to notice that the $MSE_{test}$ vary across different data splits. Hence, it is a good idea average the out-of-sample MSE over different data splits to get valid results.

Nevertheless, we observe that, based on the out-of-sample $MSE$, the basic model using OLS regression performs is about as well (or slightly better) than the flexible model.

Next, let us use lasso regression in the flexible model instead of OLS regression. Lasso (least absolute shrinkage and selection operator) is a penalized regression method that can be used to reduce the complexity of a regression model when the number of regressors $p$ is relatively large in relation to $n$.

Note that the out-of-sample $MSE$ on the test sample can be computed for any other black-box prediction method as well. Thus, let us finally compare the performance of lasso regression in the flexible model to OLS regression.

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**
```{r flexible-lass-random}
# flexible model using lasso

# estimating the parameters
reglasso <- rlasso(flex, data=train, post=FALSE)

# calculating the out-of-sample MSE
trainreglasso<- predict(reglasso, newdata=test)
MSE.lasso <- sum((y.test-trainreglasso)^2)/length(y.test)
R2.lasso<- 1- MSE.lasso/var(y.test)
```
* Test $MSE$ for the lasso on flexible model: `r MSE.lasso`

* Test $R^2$ for the lasso flexible model: `r R2.lasso`

:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

**Python code**
```{python}
# # flexible model using lasso
# # get exogenous variables from training data used in flex model
# flex_results_0 = smf.ols(flex , data=train)
# X_train = flex_results_0.exog
# print(X_train.shape)
# 
# # Get endogenous variable 
# lwage_train = train["lwage"]
# print(lwage_train.shape)
```


```{python}
# # calculating the out-of-sample MSE
# reg = linear_model.Lasso(alpha=0.1)
# lwage_lasso_fitted = reg.fit(X_train, lwage_train).predict( X_test )
# 
# MSE_lasso = np.sum((lwage_test-lwage_lasso_fitted)**2)/len(lwage_test)
# R2_lasso  = 1 - MSE_lasso/np.var(lwage_test)
# 
# print("Test MSE for the flexible model: ", MSE_lasso, " ")
# print("Test R2 for the flexible model: ", R2_lasso)
```

:::
::::::

Finally, let us summarize the results:

:::::: {.columns}
::: {.column width="48%" data-latex="{0.48\textwidth}"}

**R code**

```{r summarize-results}
Models <- c("Basic regression","Flexible regression","Lasso regression")
R_2_SAMPLE <- c(R2.test1,R2.test2,R2.lasso)
MSE_SAMPLE<- c(MSE.test1,MSE.test2,MSE.lasso)

data.frame(Models,R_2_SAMPLE,MSE_SAMPLE) %>%
  kable("markdown",caption = "Descriptive Statistics - Random Process")
```


```{r table-four}
print(data.frame(Models,R_2,MSE),type="latex")
```
:::
::: {.column width="4%" data-latex="{0.04\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::
:::::: {.column width="48%" data-latex="{0.48\textwidth}"}

**Python code**
```{python}
# Package for latex table 
# import array_to_latex as a2l
# 
# table2 = np.zeros((3, 2))
# table2[0,0] = MSE_test1
# table2[1,0] = MSE_test2
# table2[2,0] = MSE_lasso
# table2[0,1] = R2_test1
# table2[1,1] = R2_test2
# table2[2,1] = R2_lasso
# 
# table2 = pd.DataFrame(table2, columns = ["$MSE_{test}$", "$R^2_{test}$"], \
#                       index = ["basic reg","flexible reg","lasso regression"])
# table2
```

```{python}
# table2.to_latex
# print(table2.to_latex())
```

:::
::::::